---
title: "CUUPeat: Caithness and East Sutherland, Scotland - Broad-scale bare peat mapping"
date: "14/09/2020"
author: 'JNCC'
licence: 'MIT Licence'
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE,echo=F}

knitr::opts_chunk$set(echo = TRUE)
# Built with r 3.6.0
library(raster)
library(dplyr)
library(sf)
library(dplyr)
library(tmap)
library(stringr)
library(JNCCsdms)
library(tmap)
library(purrr)
library(gdalUtils)
library(sp)

#dependencies
# JNCCsdms_0.1.0    stringr_1.4.0    
# purrr_0.3.4       gdalUtils_2.0.3.2
# tmap_3.0          sf_0.9-2         
# dplyr_0.8.5       raster_3.0-7     
# sp_1.4-2

```

<div class="mycontent">

CUU Peat Project - Forest Research site - Caithness and East Sutherland, Scotland
Broad-scale bare peat mapping processing


## Site location

```{r sitemap,message=F,warning=F,eval=F}

#new S2 AOI
newAOI <- sf::st_read('Dalchork_APGB_AOI.shp') %>% dplyr::filter(id==1)
sf::st_write(newAOI, 'Dalchork_S2_AOI_new.shp')
S2 <- newAOI


#site boundaries
site <- sf::st_read('CES_Site_Boundaries.shp',quiet=T) 
site_1 <- sf::st_intersection(site,newAOI)

library(tmap)
tmap::tmap_mode("view")

tmap::tm_shape(site_1) +  tmap::tm_borders(alpha=0.5) +  tmap::tm_fill("green")+
tmap::tm_shape(S2) +  tmap::tm_borders(alpha=1,col='blue') 


```


## Preparing Variable layers

### DTM and slope

```{r dtmslopevars, eval = FALSE}


projfolder <- 'PeatlandCondition/'

#crop dtm to AOI
dtm <- raster::raster('UK_DEM_Composite_10m_WGS84.tif')
AOI= sf::st_read('Dalchork_S2_AOI_new.shp',quiet=T)

# reproject AOI
AOI_WGS84 <- AOI %>% sf::st_transform(4326)

#crop to dtm
dtm_crop <-raster::crop(dtm,AOI_WGS84)
dtm_bng <- raster::projectRaster(dtm_crop, crs=sp::CRS('+init=epsg:27700'), method='bilinear')

raster::writeRaster(dtm_bng,paste0(projfolder,'FR_CES/Variable_layers/FR_CES_dtm.tif'),overwrite=T)

# we can then create a slope layer using gdalutils
  gdalUtils::gdal_setInstallation()
  gdalUtils::gdaldem(mode="slope",
                     input_dem=paste0(projfolder,'FR_CES/Variable_layers/FR_CES_dtm.tif'),
                     output=paste0(projfolder,'FR_CES/Variable_layers/FR_CES_slope.tif'),
                    ,verbose=T)

```

### Climate layers

```{r prepclimate,eval=F}

#list all data files from met office
met_files <- data.frame(files=list.files("MetOffice_HadUK_monthly_grid_1km/",pattern='.nc'))
#subset to seasonal records
met_season <- met_files %>% dplyr::filter(stringr::str_detect(files, "Seas"))
#iterate through taking just the summer season data
out<- purrr:::map(met_season$files,function(metlayer){
  met_raw <- raster::brick(paste0("MetOffice_HadUK_monthly_grid_1km/",as.character(metlayer)))
  summer <- met_raw[[3]]
  summer_crop <-raster::crop(summer,AOI)
  #get filename of measure and year
  filename_split <- stringr::str_split(gsub(as.character(metlayer),pattern="Had",replacement="_Had"),pattern='_')
filename <- paste0(stringr::str_extract(as.character(filename_split[[1]][4]),'//d{4}'),'_',filename_split[[1]][2])
#write out
  raster::writeRaster(summer_crop, paste0(projfolder,'FR_CES/Variable_layers/FR_CES_',filename,'.tif'),overwrite=T)
  
})

#create a mean layer for use with years without data released - 2019,2020
vars <- c("rainfall","tas","tasmin","tasmax")
# iterate through each variable
all <- purrr::map(vars, .f=function(var){
  varfiles <- list.files(paste0(projfolder,'FR_CES/Variable_layers/'),full.name=T,pattern=var) # list all files in that folder for variable
  varfiles_r <- purrr::map(varfiles, raster::raster) #load
  varstack <- raster::stack(varfiles_r) #stack 
  varmean <- raster::calc(varstack,mean) #mean
  raster::writeRaster(varmean,paste0(projfolder,'FR_CES/Variable_layers/FR_CES_mean_',var,'.tif')) #write out
})


```


### Masking Sentinel imagery and calculating indices

extraction time series: 
*	2016-06-05 - S2A_20160605_lat58lon377_T30VVK_ORB123_utm30n_osgb
	
Modelling time series

* 2015	S2A_20150929_lat58lon377_T30VVK_ORB123_utm30n_osgb	2015-09-29
* 2016	S2A_20160605_lat58lon377_T30VVK_ORB123_utm30n_osgb	2016-06-05
* 2017	s3://sentinel-s2-l2a/tiles/30/V/VK/2017/9/23/0      2017-09-23
* 2018	s3://sentinel-s2-l2a/tiles/30/V/VK/2018/6/30/0	    2018-06-30
* 2019	S2A_20190826_lat58lon377_T30VVK_ORB080_utm30n_osgb	2019-08-26
* 2020	s3://sentinel-s2-l2a/tiles/30/V/VK/2020/5/07/0	    2020-05-07


```{r, eval=F}
projfolder <- 'PeatlandCondition/'

#prep and crop sentinel imagery to new AOI
s2_processing(s2path=paste0(projfolder,"Data/Sentinel-2/FR_Dalchork/"),
              out_folder = paste0(projfolder,"FR_CES/Variable_layers/"),
              nirband=8,rband=3,bband=1,gband=2,swirband=9,
              indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
              cropAOI=paste0(projfolder,'Data/Site_AOIs/FR_Dalchork/Dalchork_S2_AOI_new.shp')

```


## Compiling Variable Stacks

```{r,eval=F}
# create varstacks and mask

# extraction varstack: 2016-06-05 (same img as 2016 in time series)
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'FR_CES/Variable_layers/'),
             satimg='SEN2_20160605_lat58lon377_T30VVK_ORB123_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Dalchork_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('FR_CES_2016_rainfall.tif',
                    'FR_CES_2016_tas.tif',
                    'FR_CES_2016_tasmin.tif',
                    'FR_CES_2016_tasmax.tif',
                    'FR_CES_slope.tif'),
             mask= paste0(projfolder,'FR_CES/Shapefiles/CES_Site1_nonPlanted.shp'),
             outpath = paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/'),
             stackname='FR_CES_2016')

# 2015 varstack: 2015-09-29
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'FR_CES/Variable_layers/'),
             satimg='SEN2_20150929_lat58lon377_T30VVK_ORB123_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Dalchork_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('FR_CES_2015_rainfall.tif',
                    'FR_CES_2015_tas.tif',
                    'FR_CES_2015_tasmin.tif',
                    'FR_CES_2015_tasmax.tif',
                    'FR_CES_slope.tif'),
             mask= paste0(projfolder,'FR_CES/Shapefiles/CES_Site1_nonPlanted.shp'),
             outpath = paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/'),
             stackname='FR_CES_2015')

# 2017 varstack: 2017-09-23
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'FR_CES/Variable_layers/'),
             satimg='SEN2_20170923_lat58lon377_T30VVK_ORB123_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Dalchork_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('FR_CES_2017_rainfall.tif',
                    'FR_CES_2017_tas.tif',
                    'FR_CES_2017_tasmin.tif',
                    'FR_CES_2017_tasmax.tif',
                    'FR_CES_slope.tif'),
             mask= paste0(projfolder,'FR_CES/Shapefiles/CES_Site1_nonPlanted.shp'),
             outpath = paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/'),
             stackname='FR_CES_2017')

# 2018 varstack: 2018-06-30
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'FR_CES/Variable_layers/'),
             satimg='SEN2_20180630_lat58lon377_T30VVK_ORB123_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Dalchork_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('FR_CES_2018_rainfall.tif',
                    'FR_CES_2018_tas.tif',
                    'FR_CES_2018_tasmin.tif',
                    'FR_CES_2018_tasmax.tif',
                    'FR_CES_slope.tif'),
             mask= paste0(projfolder,'FR_CES/Shapefiles/CES_Site1_nonPlanted.shp'),
             outpath = paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/'),
             stackname='FR_CES_2018')

# 2019 varstack: 2019-08-26
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'FR_CES/Variable_layers/'),
             satimg='SEN2_20190826_lat58lon377_T30VVK_ORB080_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Dalchork_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('FR_CES_mean_rainfall.tif',
                    'FR_CES_mean_tas.tif',
                    'FR_CES_mean_tasmin.tif',
                    'FR_CES_mean_tasmax.tif',
                    'FR_CES_slope.tif'),
             mask= paste0(projfolder,'FR_CES/Shapefiles/CES_Site1_nonPlanted.shp'),
             outpath = paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/'),
             stackname='FR_CES_2019')

# 2020 varstack: 2020-05-07
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'FR_CES/Variable_layers/'),
             satimg='SEN2_20200507_lat58lon377_T30VVK_ORB080_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Dalchork_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('FR_CES_mean_rainfall.tif',
                    'FR_CES_mean_tas.tif',
                    'FR_CES_mean_tasmin.tif',
                    'FR_CES_mean_tasmax.tif',
                    'FR_CES_slope.tif'),
             mask= paste0(projfolder,'FR_CES/Shapefiles/CES_Site1_nonPlanted.shp'),
             outpath = paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/'),
             stackname='FR_CES_2020')
```

## Extracting training data

```{r,eval=F}
# For FR site only one closest date for the 2016 training data 
train_extract(varstack=paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/FR_CES_2016.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope'),
              trainr=paste0(projfolder,'FR_CES/Fine_scale_mapping/bare_pcov_mosaic.tif'),
              outfolder=paste0(projfolder,'FR_CES/Time_series_modelling/Training_data/'),
              stratified = T)

```


## Running the time series regression models


```{r,eval=F}
projfolder <- 'PeatlandCondition/'

## look at how many per category
training <- read.csv(paste0(projfolder,"FR_CES/Time_series_modelling/Training_data/FR_CES_2016.csv"))
training %>% dplyr::group_by(barecat) %>% dplyr::summarise(n=n())


## covariate analysis
varstack <- raster::stack(paste0(projfolder,'FR_CES/Time_series_modelling/Varstacks/FR_CES_2016.tif'))
names(varstack) <-c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope')
# run  function to see how the variables correlate
JNCCsdms::corrVars(varstack)


## train models and run predictions
RFReg(training=paste0(projfolder,"FR_CES/Time_series_modelling/Training_data/FR_CES_2016.csv"),
      varpredict=paste0(projfolder,"FR_CES/Time_series_modelling/Varstacks/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"FR_CES/Time_series_modelling/RegressionModels/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=10000, 
      resamp=5,
      stratified=T,fillsamp=F)

```

## Assessing change over time

```{r,eval=F}
#create folder to save outputs
dir.create(paste0(projfolder,"FR_CES/Time_series_modelling/RegressionModels/ChangeMaps"))
#list all predicted maps
pred_maps <- list.files(paste0(projfolder,"FR_CES/Time_series_modelling/RegressionModels/Outputs/"),
                        pattern='.tif',full.names = T)
#load in all maps
pred_tifs <- purrr::map(pred_maps,raster::raster)
#get year names
names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")
#make comparison table
year_combo <- data.frame(startyear=c(names(pred_tifs)[1:5],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:6],names(pred_tifs)[6]))
#iterate through comparison table calculating end-start to get change 
purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('FR_CES',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"FR_CES/Time_series_modelling/RegressionModels/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})

```
