---
title: "CUUPeat: Dark Peak, England- Fine-scale bare peat mapping"
date: "14/09/2020"
author: 'JNCC'
licence: 'MIT Licence'
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE,echo=F}

knitr::opts_chunk$set(echo = TRUE)
# Built with r 3.6.0
library(sf)
library(purrr)
library(tmap)
library(dplyr)
library(raster)
library(gdalUtils)
library(stringr)

#package dependencies
## purrr_0.3.4       tmap_3.0          sf_0.9-2          stringr_1.4.0  
## dplyr_0.8.5       raster_3.0-7      gdalUtils_2.0.3.2 
```

<div class="mycontent">

CUU Peat Project - Natural England site - Dark Peak, England
Fine-scale bare peat mapping processing

## Site location

```{r}

#site boundaries
site <- sf::st_read('NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp',quiet=T)
#S2 AOI
S2 <- sf::st_read('NE_DarkPeak/DarkPeak_S2_AOI.shp',quiet=T)
#APGB AOI
APGB <- sf::st_read('NE_DarkPeak/DarkPeak_APGB_AOI.shp',quiet=T)

library(tmap)
tmap::tmap_mode("view")

tmap::tm_shape(site) +  tmap::tm_borders(alpha=0.5) +  tm_fill("green")+
  tmap::tm_shape(S2) +  tmap::tm_borders(alpha=1,col='blue') +
  tmap::tm_shape(APGB) +  tmap::tm_borders(alpha=1,col='pink')


```


## Fine-scale map from APGB

For the Dark Peak site, there were APGB imagery available from:

* 2018-06-27

The majority of the site was previously processed during the pilot study so it was agreed to use these training data for the regression modelling and not include any further APGB imagery.


### masking classified tiles to boundaries and renaming with dates

```{r, eval=F}

## find all tiles within site
#site boundaries
site <- sf::st_read('NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp',quiet=T)
imagedate <- sf::st_read('CIR_Date_Grid.shp')

imagedat <- sf::st_intersection(imagedate,site)
x<- data.frame(unique(imagedat$OS))                 
                     
#mask to shapefiles - this was done in chunks as 564 tiles
img_fold <- 'PeatlandCondition'

MaskPeatSoils(img_folder=paste0(img_fold,'Data/APGB/NE_DarkPeak/'),
              mask=paste0(img_fold,'NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp',
              delete=T)

## rename files -  modified APGB dates function
lookup <- 'CIR_Date_Grid.shp'
imgfolder<- paste0(img_fold,'Data/APGB/NE_DarkPeak/')
field='OS'


  date_lookup <- sf::st_read(lookup,quiet=T)
  imgs <- list.files(imgfolder,pattern='tif',full.names = T)

  #get dates
  date_df <- date_lookup %>% sf::st_drop_geometry() %>% dplyr::select(field,DateFlown) %>%
    tidyr::separate(DateFlown,'date',sep=",",extra="drop")
  purrr::map(imgs,.f=function(imgname){
    date <- date_df %>% dplyr::filter(get(field) == gsub(gsub(basename(imgname),pattern='.tif',replacement=""),pattern="BARE_",replacement=""))

    file.rename(imgname, paste0(dirname(imgname),'/',as.character(date$date),'_',basename(imgname)))
    imgname
  })
  
  # remove class 2 and 3 from images ( previously in the pilot these represented rock and vegetation)
  bare_path <- paste0(dirpath,'Fine_scale_mapping/bare/')
  imgs <- list.files(bare_path,pattern='tif',full.names=T)
  purrr::map(imgs, .f=function(tile){
    tile_bare <- raster::raster(tile)
    tile_bare[tile_bare==2 |tile_bare==3]<-0
    raster::writeRaster(tile_bare,tile,overwrite=T)
  })

```

At this stage, the pilot data is already up to the classed stage 

```{r,eval=F}
dirpath <- 'NE_DarkPeak/'

 #### Converting to percentage cover of bare peat at 10m pixel ####
bare_path <- paste0(dirpath,'Fine_scale_mapping/bare/')
out_path <- paste0(dirpath,'Fine_scale_mapping/')
polymask <- paste0(dirpath,'Fine_scale_mapping/Shapefiles/DarkPeak_Site_Boundaries.shp')

#list all files
  imgs <- list.files(bare_path,pattern='tif',full.names=T)
  #create folders
  if(!dir.exists(paste0(out_path,'bare_pcov'))){
    dir.create(paste0(out_path,'bare_pcov'))
  }

  habmask <- sf::st_read(polymask)
  
  #iterate through bare classed tiles
  purrr::map(imgs, .f=function(tile){
    tile_bare <- raster::raster(tile)
    #get tile name
    name <- gsub(unlist(stringr::str_split(basename(tile),'_'))[3],pattern='.tif',replacement='')
    namenobare <- gsub(basename(tile),pattern='BARE_',replacement='')

    #convert NA values to 0
    tile_bare[is.na(tile_bare)]<-0
    raster::writeRaster(tile_bare,paste0(out_path,'bare_pcov/baretmp_',basename(tile)),overwrite=T)
    #create a 10m pixel version of raster
    gdalUtils::gdal_setInstallation()
    gdalUtils::gdalwarp(srcfile = paste0(out_path,'bare_pcov/baretmp_',basename(tile)),
                        dstfile = paste0(out_path,'bare_pcov/pcov_',basename(tile)),
                        s_srs = 'EPSG:27700',t_srs = 'EPSG:27700',
                        r = 'average',tr=c(10,10), overwrite = T, verbose = T)
    #annoyingly gdalwarp doesnt summarise on sum so we will use average*pixelcellcount = sum
    ## calculate no.cells in high res compared to low res
    highres_count <- (10 / res(tile_bare)[1])^2

    pcov_img <- raster::raster(paste0(out_path,'bare_pcov/pcov_',basename(tile)))
    pcov_df <- raster::as.data.frame(pcov_img)
    names(pcov_df) <- 'average'
    pcov_df <- pcov_df %>% dplyr::mutate(sum = average*highres_count) %>%
      dplyr::mutate(pcov = sum/highres_count)
    pcov_r <- raster::setValues(pcov_img,values=pcov_df$pcov)

    #mask to remove border effect of incomplete cells
    pcov_mask <- raster::mask(pcov_r,habmask)

    writeRaster(pcov_mask,paste0(out_path,'bare_pcov/pcov_',namenobare),
                overwrite=T)
    file.remove(paste0(out_path,'bare_pcov/baretmp_',basename(tile)))
    print(paste0(name, ' done.'))

    })

  #mosaic together
  pcov <- list.files(paste0(dirpath,'Fine_scale_mapping/bare_pcov/'),full.names=T)
  pcov_list <- purrr::map(pcov,raster::raster)
  names(pcov_list) <- NULL
  pcov_list$fun <- mean
  pcov_mosaic <- do.call(raster::mosaic,c(pcov_list,progress="window"))
  raster::writeRaster(pcov_mosaic,paste0(dirpath,'Fine_scale_mapping/bare_pcov_mosaic.tif'),overwrite=T)
                    
                    
```









