---
title: "CUUPeat: Dark Peak - Broad-scale bare peat mapping"
date: "14/09/2020"
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE,echo=F}

knitr::opts_chunk$set(echo = TRUE)
# Built with r 3.6.0
library(tmap)
library(raster)
library(dplyr)
library(sf)
library(stringr)
library(gdalUtils)
library(purrr)
library(JNCCsdms)

#package dependencies
# JNCCsdms_0.1.0    stringr_1.4.0    
# purrr_0.3.4       gdalUtils_2.0.3.2
# tmap_3.0          sf_0.9-2         
# dplyr_0.8.5       raster_3.0-7     
# sp_1.4-2 

```

<div class="mycontent">

CUU Peat Project - Natural England site - Dark Peak, England


## Site location

```{r sitemap,eval=F}
#site boundaries
site <- sf::st_read('NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp',quiet=T)
#S2 AOI
S2 <- sf::st_read('NE_DarkPeak/DarkPeak_S2_AOI.shp',quiet=T)
#APGB AOI
APGB <- sf::st_read('NE_DarkPeak/DarkPeak_APGB_AOI.shp',quiet=T)

#plot
tmap::tmap_mode("view")
tmap::tm_shape(site) +  tmap::tm_borders(alpha=0.5) +  tmap::tm_fill("green") +
  tmap::tm_shape(S2) +  tmap::tm_borders(alpha=1,col='blue') +
  tmap::tm_shape(APGB) +  tmap::tm_borders(alpha=1,col='pink')


```

## Preparing Variable layers

### DTM and slope


```{r dtmslopevars, eval = FALSE}

projfolder <- 'PeatlandCondition/'

#crop dtm to AOI
dtm <- raster::raster('UK_DEM_Composite_10m_WGS84.tif')
AOI= sf::st_read(paste0(projfolder,'Data/Site_AOIs/NE_DarkPeak/DarkPeak_S2_AOI.shp'),quiet=T)

# reproject AOI
AOI_WGS84 <- AOI %>% sf::st_transform(4326)

#crop to dtm
dtm_crop <-raster::crop(dtm,AOI_WGS84)
dtm_bng <- raster::projectRaster(dtm_crop, crs=sp::CRS('+init=epsg:27700'), method='bilinear')
raster::writeRaster(dtm_bng,paste0(projfolder,'NE_DarkPeak/Variable_layers/NE_DP_dtm.tif'),overwrite=T)

# we can then create a slope layer using gdalutils
gdalUtils::gdal_setInstallation()
gdalUtils::gdaldem(mode="slope",
                     input_dem=paste0(projfolder,'NE_DarkPeak/Variable_layers/NE_DP_dtm.tif'),
                     output=paste0(projfolder,'NE_DarkPeak/Variable_layers/NE_DP_slope.tif'),
                    ,verbose=T)

```



### Climate layers

```{r prepclimate,eval=F}
#list all data files from met office
met_files <- data.frame(files=list.files("MetOffice_HadUK_monthly_grid_1km/",pattern='.nc'))
#subset to seasonal records
met_season <- met_files %>% dplyr::filter(stringr::str_detect(files, "Seas"))
#iterate through taking just the summer season data
out<- purrr:::map(met_season$files,function(metlayer){
  met_raw <- raster::brick(paste0("MetOffice_HadUK_monthly_grid_1km/",as.character(metlayer)))
  summer <- met_raw[[3]]
  summer_crop <-raster::crop(summer,AOI)
  #get filename of measure and year
  filename_split <- stringr::str_split(gsub(as.character(metlayer),pattern="Had",replacement="_Had"),pattern='_')
filename <- paste0(stringr::str_extract(as.character(filename_split[[1]][4]),'\\d{4}'),'_',filename_split[[1]][2])
#write out
  raster::writeRaster(summer_crop, paste0(projfolder,'NE_DarkPeak/Variable_layers/NE_DP_',filename,'.tif'),overwrite=T)
  summer_crop
})



#create a mean layer for use with years without data released - 2019,2020
vars <- c("rainfall","tas","tasmin","tasmax")
# iterate through each variable
all <- purrr::map(vars, .f=function(var){
  varfiles <- list.files(paste0(projfolder,'NE_DarkPeak/Variable_layers/'),full.name=T,pattern=var) # list all files in that folder for variable
  varfiles_r <- purrr::map(varfiles, raster::raster) #load
  varstack <- raster::stack(varfiles_r) #stack 
  varmean <- raster::calc(varstack,mean) #mean
  raster::writeRaster(varmean,paste0(projfolder,'NE_DarkPeak/Variable_layers/NE_DP_mean_',var,'.tif')) #write out
})

```

### Sentinel imagery and calculating indices

extraction time series: 
* S2A_20180629_lat54lon217_T30UWE_ORB037_utm30n_osgb	2018-06-29

Modelling time series:
* 2016	 S2A_20160719_lat54lon217_T30UWE_ORB037_utm30n_osgb	2016-07-19
* 2017	S2B_20171027_lat54lon217_T30UWE_ORB037_utm30n_osgb	2017-10-27
* 2018	S2A_20180629_lat54lon217_T30UWE_ORB037_utm30n_osgb	2018-06-29
* 2019	S2A_20190422_lat54lon217_T30UWE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref	2019-04-22
*  2020	S2A_20200625_lat54lon217_T30UWE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref	2020-06-25


```{r,eval=F}
projfolder <- 'PeatlandCondition/'

#prep S2 imagery  
s2_processing(s2path=paste0(projfolder,"Data/Sentinel-2/NE_DarkPeak/"),
              out_folder = paste0(projfolder,"NE_DarkPeak/Variable_layers/"),
              nirband=8,rband=3,bband=1,gband=2,swirband=9,
              indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
              cropAOI=NA)

```

## Compiling Variable Stacks

```{r,eval=F}
# create varstacks and mask

# extraction varstack: 2018-06-29 (same img as 2018 in time series)
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'NE_DarkPeak/Variable_layers/'),
             satimg='SEN2_20180629_lat54lon217_T30UWE_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_DarkPeak_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_DP_2018_rainfall.tif',
                    'NE_DP_2018_tas.tif',
                    'NE_DP_2018_tasmin.tif',
                    'NE_DP_2018_tasmax.tif',
                    'NE_DP_slope.tif'),
             mask= paste0(projfolder,'NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp'),
             outpath = paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Varstacks/'),
             stackname='NE_DP_2018')

# 2016 varstack: 2016-07-19
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'NE_DarkPeak/Variable_layers/'),
             satimg='SEN2_20160719_lat54lon217_T30UWE_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_DarkPeak_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_DP_2016_rainfall.tif',
                    'NE_DP_2016_tas.tif',
                    'NE_DP_2016_tasmin.tif',
                    'NE_DP_2016_tasmax.tif',
                    'NE_DP_slope.tif'),
             mask= paste0(projfolder,'NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp'),
             outpath = paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Varstacks/'),
             stackname='NE_DP_2016')

# 2017 varstack: 2017-10-27
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'NE_DarkPeak/Variable_layers/'),
             satimg='SEN2_20171029_lat54lon217_T30UWE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_DarkPeak_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_DP_2017_rainfall.tif',
                    'NE_DP_2017_tas.tif',
                    'NE_DP_2017_tasmin.tif',
                    'NE_DP_2017_tasmax.tif',
                    'NE_DP_slope.tif'),
             mask= paste0(projfolder,'NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp'),
             outpath = paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Varstacks/'),
             stackname='NE_DP_2017')

# 2019 varstack: 2019-04-22
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'NE_DarkPeak/Variable_layers/'),
             satimg='SEN2_20190422_lat54lon217_T30UWE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_DarkPeak_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_DP_mean_rainfall.tif',
                    'NE_DP_mean_tas.tif',
                    'NE_DP_mean_tasmin.tif',
                    'NE_DP_mean_tasmax.tif',
                    'NE_DP_slope.tif'),
             mask= paste0(projfolder,'NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp'),
             outpath = paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Varstacks/'),
             stackname='NE_DP_2019')

# 2020 varstack: 2020-06-25
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip
sat_varstack(varpath = paste0(projfolder,'NE_DarkPeak/Variable_layers/'),
             satimg='SEN2_20200625_lat54lon217_T30UWE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_DarkPeak_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_DP_mean_rainfall.tif',
                    'NE_DP_mean_tas.tif',
                    'NE_DP_mean_tasmin.tif',
                    'NE_DP_mean_tasmax.tif',
                    'NE_DP_slope.tif'),
             mask= paste0(projfolder,'NE_DarkPeak/Shapefiles/DarkPeak_Site_Boundaries.shp'),
             outpath = paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Varstacks/'),
             stackname='NE_DP_2020')

```


## Extracting training data

```{r,eval=F}
# For DP site there was only one year of training data in 2018
train_extract(varstack=paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Varstacks/NE_DP_2018.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
              trainr=paste0(projfolder,'NE_DarkPeak/Fine_scale_mapping/bare_pcov_mosaic.tif'),
              outfolder=paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Training_data/'),
              stratified = T)

```

## Running the time series regression models
```{r,eval=F}
projfolder <- 'PeatlandCondition/'

## look at how many per category
training <- read.csv(paste0(projfolder,"NE_DarkPeak/Time_series_modelling/Training_data/NE_DP_2018.csv"))
training %>% dplyr::group_by(barecat) %>% dplyr::summarise(n=n())


## covariate analysis
varstack <- raster::stack(paste0(projfolder,'NE_DarkPeak/Time_series_modelling/Varstacks/FR_CES_2016.tif'))
names(varstack) <-c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope')
# run  function to see how the variables correlate
JNCCsdms::corrVars(varstack)


## train models and run predictions
RFReg(training=paste0(projfolder,"NE_DarkPeak/Time_series_modelling/Training_data/NE_DP_2018.csv"),
      varpredict=paste0(projfolder,"NE_DarkPeak/Time_series_modelling/Varstacks/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"NE_DarkPeak/Time_series_modelling/RegressionModels/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=10000, 
      resamp=5,
      stratified=T,fillsamp=F)

```

## Assessing change over time

```{r,eval=F}
dir.create(paste0(projfolder,"NE_DarkPeak/Time_series_modelling/RegressionModels/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"NE_DarkPeak/Time_series_modelling/RegressionModels/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:4],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:5],names(pred_tifs)[5]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('NE_DarkPeak',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"NE_DarkPeak/Time_series_modelling/RegressionModels/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})

```
