---
title: "CUUPeat: Humberhead NNR - Fine-scale bare peat mapping"
date: "15/09/2020"
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE,echo=F}

knitr::opts_chunk$set(echo = TRUE)
# Built with r 3.6.0
library(sf)
library(purrr)
library(tmap)
library(dplyr)
library(raster)

#package dependencies
## purrr_0.3.4       tmap_3.0          sf_0.9-2         
## dplyr_0.8.5       raster_3.0-7  stringr_1.4.0
## tictoc_1.0        lubridate_1.7.4 magrittr_1.5

```

<div class="mycontent">

CUU Peat Project - Natural England site - Humberhead NNR
Fine-scale bare peat mapping processing

## Site location

```{r sitemap,message=F,warning=F}

#site boundaries
site <- sf::st_read('Humberhead_Site_Boundaries.shp',quiet=T)
#S2 AOI
S2 <- sf::st_read('ThorneandHatfield_S2_AOI.shp',quiet=T)
#APGB AOI
APGB <- sf::st_read('ThorneandHatfield_APGB_AOI.shp',quiet=T)

tmap::tmap_mode("view")

tmap::tm_shape(site) +  tmap::tm_borders(alpha=0.5) +  tmap::tm_fill("green") +
  tmap::tm_shape(S2) +  tmap::tm_borders(alpha=1,col='blue') +
  tmap::tm_shape(APGB) +  tmap::tm_borders(alpha=1,col='pink')


```

## Preparing the APGB

For the Humberhead site, there were a mixture of APGB imageery available from:

Thorne:

* 2016-04-20
* 2016-04-21
* 2017-05-25
* 2017-05-25
* 2019-10-20

Hatfield:

* 2016-04-21
* 2016-05-05
* 2018-07-01
* 2019-10-20

Preparing the latest site APGB imagery:
```{r apgb_latest, include=T,message=F,warning=F,fig.show = "hold", out.width = "50%", fig.align = "default",eval=F}
dirpath <- "PeatlandCondition/Data/APGB/WG_BB/"

#run function to resample nir band from image b, and combine with image a. the nir argument specifys which band is the nir band in image b and the start and end arguments are for iterating between many data files. 

AOIs <- list.files(paste0(dirpath,"25cm Aerial Photo/"))
purrr::map(AOIs,.f=function(shape){
  comboAPGB(a.path=paste0(dirpath,"25cm Aerial Photo/",shape,"/"),
          b.path=paste0(dirpath,"50cm Colour Infrared/",gsub(shape,pattern="1_RGB",replacement="2_CIR"),"/"),
          nir=1,
          out.path=paste0(dirpath),
          start=1,end=NULL)
})

## rename files with dates
date_lookup <- 'CIR_Date_Grid.shp'

APGBdate(imgfolder=paste0(dirpath,'combined/Shape1/'),lookup=date_lookup)
APGBdate(imgfolder=paste0(dirpath,'combined/Shape2/'),lookup=date_lookup)


```

Preparing the second year of APGB imagery available for the site (partial coverage):
```{r apgb_archive, include=T,message=F,warning=F,fig.show = "hold", out.width = "50%", fig.align = "default",eval=F}

## Processing the latest APGB
dirpath <- "PeatlandCondition/Data/APGB/NE_HumberheadNNR/"

#run function to resample nir band from image b, and combine with image a. the nir argument specifys which band is the nir band in image b and the start and end arguments are for iterating between many data files. 

AOIs <- list.files(paste0(dirpath,"25cm Aerial Photo/"))
purrr::map(AOIs[3],.f=function(shape){
  comboAPGB(a.path=paste0(dirpath,"25cm Aerial Photo/",shape,"/"),
          b.path=paste0(dirpath,"50cm Colour Infrared/",gsub(shape,pattern="3_RGB",replacement="4_CIR"),"/"),
          nir=1,
          out.path=paste0(dirpath),
          start=1,end=NULL)
})

#rename files
date_lookup <- '111818-2_CIR_1_Shape_Date_Grid.shp'

APGBdate(imgfolder=paste0(dirpath,'combined/111818_Shape1/'),
         lookup=date_lookup,field='kmRef')
APGBdate(imgfolder=paste0(dirpath,'combined/111818_Shape2/'),
         lookup=date_lookup,field='kmRef')

```

## Creating fine scale maps

### Mask APGB imagery to peat soils and remove any tiles not over peat soils

```{r,eval=F }

#site boundaries
site <- sf::st_read('Humberhead_Site_Boundaries.shp',quiet=T)

# copy folder contents
#Thorne
file.copy("Data/APGB/NE_HumberheadNNR/Combined/Shape1/",
          "NE_HumberheadNNR/Fine_scale_mapping/Thorne/Masked",recursive=T)
file.copy("Data/APGB/NE_HumberheadNNR/Combined/11818_Shape1/",
          "NE_HumberheadNNR/Fine_scale_mapping/Thorne/Masked",recursive=T)
#Hatfield
file.copy("Data/APGB/NE_HumberheadNNR/Combined/Shape2/",
          "NE_HumberheadNNR/Fine_scale_mapping/Hatfield/Masked",recursive=T)
file.copy("Data/APGB/NE_HumberheadNNR/Combined/11818_Shape2/",
          "NE_HumberheadNNR/Fine_scale_mapping/Hatfield/Masked",recursive=T)

#mask to the site boundaries 
img_folder="NE_HumberheadNNR/Fine_scale_mapping/Thorne/Masked/Shape1/"
#img_folder="NE_HumberheadNNR/Fine_scale_mapping/Thorne/Masked/11818_Shape1/"
#img_folder="NE_HumberheadNNR/Fine_scale_mapping/Hatfield/Masked/Shape2/"
#img_folder="NE_HumberheadNNR/Fine_scale_mapping/Hatfield/Masked/11818_Shape2/"
# masking function
MaskPeatSoils(img_folder=img_folder,
              mask='Humberhead_Site_Boundaries.shp',
              delete=T)
```

## Thorne Moors ##

Very difficult varied vegetation, small patches of obvious bare peat then lots of scrubby areas. 

## 2019 imagery
### Testing out threshold rules
```{r, eval=F}
## test out a thresholding rule on a tile 2016-06-06_NC5620
img_folder="NE_HumberheadNNR/Fine_scale_mapping/Thorne/Masked/2019/"
#load in layers
img <- '2019-10-20_SE7218'
#create indices layer
runIndices(imagepath=paste0(img_folder,img,'.tif'),
            outpath=img_folder,
            nir=4,r=1, b=3, g=2,
            indices=c("Brightness","GLI","NDVI","RG", "RB"), nf=T)
#indices
indices <- c(paste0(img_folder,"Indices/", img, "/",img,"_Brightness.tif"),
             paste0(img_folder,"Indices/", img, "/",img,"_GLI.tif"),
             paste0(img_folder,"Indices/", img, "/",img,"_NDVI.tif"),
             paste0(img_folder,"Indices/", img, "/",img,"_RG.tif"),
             paste0(img_folder,"Indices/", img, "/",img,"_RB.tif"))
indlist <- purrr::map(indices,.f=raster::raster)


nir <- raster::raster(paste0(img_folder,img,'.tif'), band=4)
indlist <- append(indlist,nir)
layer_stack <- raster::stack(indlist)


#x1 = Brightness, x2 = GLI, x3 = NDVI, x4 = RG, x5=RB 
thresh_fun <- function(x1,x2,x3,x4,x5,x6) {
  ifelse(x1 >30 & x1 <50 & x2 < 0 & x3 <0.5 & 
           x4> 0.95 &
           x5 < 1.4 & x5> 0.86 & 
           x6<120, 1, NA) 
}

#Apply function to raster stack
r.class <- raster::overlay(layer_stack, fun=thresh_fun)

if(!dir.exists(paste0(img_folder,"bare"))){
  dir.create(paste0(img_folder,"bare"))
}
raster::writeRaster(r.class,paste0(img_folder,"bare/",img,"_Bare.tif"),overwrite=T)
```

## 2017 imagery - all made using peatpal
Found threshold rules were on a tile by tile basis and quite different so instead manually identified in qgis. Several tiles were without any bare peat, therefore a mosaic was first created and then bare peat tiles were used to update this to generate the bare peat layer.

```{r, eval= F}
img_folder="NE_HumberheadNNR/Fine_scale_mapping/Thorne/"
mask<- sf::st_read('NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp')
#create mosaic layer of all tiles with nas
alltiles <- list.files(paste0(img_folder,'Masked/2017'),pattern='.tif',full.names=T)
tiles_r <- purrr::map(alltiles,raster,band=1)
names(tiles_r) <- NULL
tiles_r$fun <- mean
rast.mosaic <- do.call(raster::mosaic,c(tiles_r,progress="window"))
#make all values na
rast.mosaic[!is.na(rast.mosaic)]<- NA

# mosaic bare peat tiles
baretiles <- list.files(paste0(img_folder,'Bare/2017/'),pattern='.tif',full.names=T)
bare_r <- purrr::map(baretiles,.f=function(tif){
  baretile <- raster::raster(tif)
  baretile[is.na(baretile)]<-0
  baretile
})
names(bare_r) <- NULL
bare_r$fun <- mean
bare_mosaic <- do.call(raster::mosaic,c(bare_r,progress="window"))
bare_extended <- raster::extend(bare_mosaic,extent(rast.mosaic),value=0)
bare_extended[is.na(bare_extended)]<-0
bare_masked <- raster::mask(bare_extended,mask)

#write out layer
raster::writeRaster(bare_masked,paste0(img_folder,'Bare/2017/Humberhead_Thorne_mosaic_2017.tif',overwrite=T)
                    
                    

```

## 2016 imagery

```{r }
x<-data.frame(list.files('Data/APGB/NE_HumberheadNNR/combined/111818_Shape1 - Copy'))

```

Finally all the tiles converted in a binary classification; 0 if there was no bare peat present and 1 if there was bare peat present. This was converted to the 10m pixel level as a percentage cover per 10m of bare peat.

```{r, eval=FALSE}

options("rgdal_show_exportToProj4_warnings"="none")
source("pcovClassify.R")

tictoc::tic()
#convert into percentage covers
pcov_classify(img_path=paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Thorne/Masked/2019/'), 
              bare_path=paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Thorne/Bare/2019/'),
              out_path=paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Thorne/'),
              polymask=paste0(dirpath,'NE_HumberheadNNR/Shapefile/Humberhead_Site_Boundaries.shp'))

# mosaic tiles
pcov <- list.files(paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Thorne/bare_pcov/'),full.names=T)
pcov_list <- purrr::map(pcov,raster)
names(pcov_list)<-NULL
pcov_list$fun <- mean
pcov_mosaic <- do.call(raster::mosaic,c(pcov_list,progress="window"))
raster::writeRaster(pcov_mosaic,paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Thorne/bare_pcov_mosaic.tif'))

# finished message
sink(fs::path(dirpath, 'NE_HumberheadNNR','Fine_scale_mapping', 'Thorne', 'FINISHED.txt'))
lubridate::now()
tictoc::toc()
sink()

```

## Hatfield Moors ##

## 2016 imagery

### Trying out blanket rule on all tiles

```{r, eval=F}
#x1=Brightness, x2=GLI,x3=NDVI,x4 = NDWI, x4=RG
img_folder <- 'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/2016/'
thresh_fun <- function(x1,x2,x3,x4,x5){
  ifelse(x1 > 50 & 
           x1 <80 & 
           x2 < 0.08 & 
           x3 < 0.5 & 
           x4 < (-0.22) & 
           x5 >0.85 , 1, 0) 
}

#run thresholding rule
barethresh(Img.path=paste0(img_folder,'2016/Masked/'), 
           out.path=paste0(img_folder,'2016/'),
           spec.bands=NA, #if spectral bands are included in the thresholding, the band number
           ind.name=c("Brightness","GLI", "NDVI","NDWI","RG"), #name of indices, should be in alphabetical order and will be used in this order in the thresholding function
           c.fun=thresh_fun, #thresholding function
           nir=4, r=1,g=2,b=3, #band numbers within the imagery
           start=1)
```

### 2018 imagery

```{r, eval=F}
img_folder <- 'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/2018/'
#x1=NIR, x2=Brightness, x3=GLI,x4=NDVI,x5 = NDWI, x6= RB, x7=RG
thresh_fun <- function(x1,x2,x3,x4,x5,x6,x7){
  ifelse(x1 > 50 & 
           x1 < 100 & 
           x2 > 50 & 
           x2 < 100 &
           x3 > 0 & 
           x3 < 0.1 &
           x4 > (-0.15) & 
           x4 < 0.12 &
           x5 > (-0.1) &
           x5 < 0.15 &
           x6 > 1 &
           x6 < 1.5 &
           x7 > 0.8 &
           x7 < 1.1
         , 1, 0) 
}
```
 
```{r, eval=F}
#run thresholding rule
barethresh(Img.path=img_folder, 
           out.path=img_folder,
           spec.bands=4, #if spectral bands are included in the thresholding, the band number
           ind.name=c("Brightness","GLI", "NDVI","NDWI","RB","RG"), #name of indices, should be in alphabetical order and will be used in this order in the thresholding function
           c.fun=thresh_fun, #thresholding function
           nir=4, r=1,g=2,b=3, #band numbers within the imagery
           start=1)
```

###  2019 imagery

```{r, eval=F}
img_folder <- 'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/2019/'

#x1=Brightness, x2=GLI,x3=NDVI,x4 = NDWI, x5= RB, x6=RG
thresh_fun <- function(x1,x2,x3,x4,x5,x6){
  ifelse(x1 > 30 & 
           x1 <50 & 
           x2 < 0.01 & 
           x3 < 0.6 & 
           x4 < 0.4 & 
           x5 <1.3 &
           x6>0.9
        , 1, 0) 
}

#run thresholding rule
barethresh(Img.path=paste0(img_folder,'Masked/'), 
           out.path=paste0(img_folder,'bare/'),
           spec.bands=NA, #if spectral bands are included in the thresholding, the band number
           ind.name=c("Brightness","GLI", "NDVI","NDWI","RB","RG"), #name of indices, should be in alphabetical order and will be used in this order in the thresholding function
           c.fun=thresh_fun, #thresholding function
           nir=4, r=1,g=2,b=3, #band numbers within the imagery
           start=1)
```

Finally all the tiles converted in a binary classification; 0 if there was no bare peat present and 1 if there was bare peat present. This was converted to the 10m pixel level as a percentage cover per 10m of bare peat.

```{r, eval=FALSE}

options("rgdal_show_exportToProj4_warnings"="none")
source("pcovClassify.R")
dirpath <- 'PeatlandCondition/'

tictoc::tic()
#convert into percentage covers
pcov_classify(img_path=paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/Masked/2019/'), 
              bare_path=paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/Bare/2019/'),
              out_path=paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/'),
              polymask=paste0(dirpath,'NE_HumberheadNNR/Shapefile/Humberhead_Site_Boundaries.shp'))

# mosaic tiles
pcov <- list.files(paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/bare_pcov/'),full.names=T)
pcov_list <- purrr::map(pcov,raster)
names(pcov_list)<-NULL
pcov_list$fun <- mean
pcov_mosaic <- do.call(raster::mosaic,c(pcov_list,progress="window"))
raster::writeRaster(pcov_mosaic,paste0(dirpath,'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/bare_pcov_mosaic.tif'))

# finished message
sink(fs::path(dirpath, 'NE_HumberheadNNR','Fine_scale_mapping', 'Hatfield', 'FINISHED.txt'))
lubridate::now()
tictoc::toc()
sink()
```




