---
title: "CUUPeat: Humberhead NNR - Broad-scale bare peat mapping"
date: "15/09/2020"
author: 'JNCC'
licence: 'MIT Licence'
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE,echo=F}
knitr::opts_chunk$set(echo = TRUE)
# Built with r 3.6.0
library(raster)
library(dplyr)
library(sf)
library(dplyr)
library(tmap)
library(stringr)
library(JNCCsdms)
library(tmap)
library(purrr)
library(gdalUtils)
library(sp)

#dependencies
# tictoc_1.0        lubridate_1.7.4  
# magrittr_1.5      gdalUtils_2.0.3.2
# purrr_0.3.4       JNCCsdms_0.1.0   
# stringr_1.4.0     tmap_3.0         
# sf_0.9-2          dplyr_0.8.5      
# raster_3.0-7      sp_1.4-2 


```

<div class="mycontent">

CUU Peat Project - Natural England site - Humberhead NNR
Broad scale bare peat mapping processing

## Site location

```{r sitemap,message=F,warning=F}

#site boundaries
site <- sf::st_read('Humberhead_Site_Boundaries.shp',quiet=T)
#S2 AOI
S2 <- sf::st_read('ThorneandHatfield_S2_AOI.shp',quiet=T)
#APGB AOI
APGB <- sf::st_read('ThorneandHatfield_APGB_AOI.shp',quiet=T)

tmap::tmap_mode("view")

tmap::tm_shape(site) +  tmap::tm_borders(alpha=0.5) +  tmap::tm_fill("green") +
  tmap::tm_shape(S2) +  tmap::tm_borders(alpha=1,col='blue') +
  tmap::tm_shape(APGB) +  tmap::tm_borders(alpha=1,col='pink')


```

## Preparing Variable layers

### DTM and slope

```{r dtmslopevars, eval = FALSE}

projfolder <- 'PeatlandCondition/'

#crop dtm to AOI
dtm <- raster::raster('UK_DEM_Composite_10m_WGS84.tif')
AOI= sf::st_read(paste0(projfolder,'Data/Site_AOIs/NE_ThorneHatfield/ThorneandHatfield_S2_AOI.shp'),quiet=T)

# reproject AOI
AOI_WGS84 <- AOI %>% sf::st_transform(4326)

#crop to dtm
dtm_crop <-raster::crop(dtm,AOI_WGS84)
dtm_bng <- raster::projectRaster(dtm_crop, crs=sp::CRS('+init=epsg:27700'), method='bilinear')
raster::writeRaster(dtm_bng,paste0(projfolder,'NE_HumberheadNNR/Variable_layers/NE_HNNR_dtm.tif'),overwrite=T)

# we can then create a slope layer using gdalutils
gdalUtils::gdal_setInstallation()
gdalUtils::gdaldem(mode="slope",
                     input_dem=paste0(projfolder,'NE_HumberheadNNR/Variable_layers/NE_HNNR_dtm.tif'),
                     output=paste0(projfolder,'NE_HumberheadNNR/Variable_layers/NE_HNNR_slope.tif'),
                    ,verbose=T)

```

### Climate layers

```{r prepclimate,eval=F}
#list all data files from met office
met_files <- data.frame(files=list.files("MetOffice_HadUK_monthly_grid_1km/",pattern='.nc'))
#subset to seasonal records
met_season <- met_files %>% dplyr::filter(stringr::str_detect(files, "Seas"))
#iterate through taking just the summer season data
out<- purrr:::map(met_season$files,function(metlayer){
  met_raw <- raster::brick(paste0("MetOffice_HadUK_monthly_grid_1km/",as.character(metlayer)))
  summer <- met_raw[[3]]
  summer_crop <-raster::crop(summer,AOI)
  #get filename of measure and year
  filename_split <- stringr::str_split(gsub(as.character(metlayer),pattern="Had",replacement="_Had"),pattern='_')
filename <- paste0(stringr::str_extract(as.character(filename_split[[1]][4]),'\\d{4}'),'_',filename_split[[1]][2])
#write out
  raster::writeRaster(summer_crop, paste0(projfolder,'NE_HumberheadNNR/Variable_layers/NE_HNNR_',filename,'.tif'),overwrite=T)
  summer_crop
})

#create a mean layer for use with years without data released - 2019,2020
vars <- c("rainfall","tas","tasmin","tasmax")
# iterate through each variable
all <- purrr::map(vars, .f=function(var){
  varfiles <- list.files(paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),full.name=T,pattern=var) # list all files in that folder for variable
  varfiles_r <- purrr::map(varfiles, raster::raster) #load
  varstack <- raster::stack(varfiles_r) #stack 
  varmean <- raster::calc(varstack,mean) #mean
  raster::writeRaster(varmean,paste0(projfolder,'NE_HumberheadNNR/Variable_layers/NE_HNNR_mean_',var,'.tif')) #write out
})


```

### Sentinel imagery and calculating indices

extraction time series: 
* S2A_20160420_lat54lon066_T30UXE_ORB037_utm30n_osgb	2016-04-20
* S2A_20170505_lat54lon066_T30UXE_ORB037_utm30n_osgb	2017-05-05
* s3://sentinel-s2-l2a/tiles/30/U/XE/2018/7/1/0	      2018-07-01

Modelling time series:
* 2015	S2A_20150930_lat54lon066_T30UXE_ORB137_utm30n_osgb	2015-09-30
* 2016	S2A_20160719_lat54lon066_T30UXE_ORB037_utm30n_osgb	2016-07-19
* 2017	S2A_20170505_lat54lon066_T30UXE_ORB037_utm30n_osgb	2017-05-05
* 2018	s3://sentinel-s2-l2a/tiles/30/U/XE/2018/7/1/0	      2018-07-01
* 2019	S2A_20190823_lat54lon066_T30UXE_ORB037_utm30n_osgb	2019-08-23
* 2020	S2A_20200625_lat54lon066_T30UXE_ORB137_utm30n_osgb	2020-06-25

```{r, eval=F}

options("rgdal_show_exportToProj4_warnings"="none")
source("s2_processing.R")
source("FunctionRunIndices.R")
projfolder <- 'PeatlandCondition/'

#prep and crop sentinel imagery to new AOI
tictoc::tic()
s2_processing(s2path=paste0(projfolder, "Data/Sentinel-2/NE_ThorneHatfield/"),
              out_folder = paste0(projfolder, "NE_HumberheadNNR/Variable_layers/"),
              nirband=8,rband=3,bband=1,gband=2,swirband=9,
              indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
              cropAOI=paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'))

# finished message
sink(fs::path(projfolder, 'FINISHED.txt'))
lubridate::now()
tictoc::toc()
sink()

```

## Compiling Variable Stacks

```{r,eval=F}
library(raster)
options("rgdal_show_exportToProj4_warnings"="none")
source("sat_varstack.R")

projfolder <- 'PeatlandCondition/'

# create varstacks and mask
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope

tictoc::tic()
# Time series

# extraction time series: varstack: 2016-04-20
sat_varstack(varpath = paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),
             satimg='SEN2_20160420_lat54lon066_T30UXE_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Thorne_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_HNNR_2016_rainfall.tif',
                    'NE_HNNR_2016_tas.tif',
                    'NE_HNNR_2016_tasmin.tif',
                    'NE_HNNR_2016_tasmax.tif',
                    'NE_HNNR_slope.tif'),
             mask= paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'),
             outpath = paste0(projfolder, 'NE_HumberheadNNR/Time_series_modelling/Varstacks/'),
             stackname='NE_HNNR_2016_04')

# extraction and modelling time series: 2017-05-05 
sat_varstack(varpath = paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),
             satimg='SEN2_20170505_lat54lon066_T30UXE_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Thorne_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_HNNR_2017_rainfall.tif',
                    'NE_HNNR_2017_tas.tif',
                    'NE_HNNR_2017_tasmin.tif',
                    'NE_HNNR_2017_tasmax.tif',
                    'NE_HNNR_slope.tif'),
             mask= paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'),
             outpath = paste0(projfolder, 'NE_HumberheadNNR/Time_series_modelling/Varstacks/'),
             stackname='NE_HNNR_2017')


# extraction and modelling time series: 2018-07-01
sat_varstack(varpath = paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),
             satimg='SEN2_20180701_lat54lon066_T30UXE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Thorne_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_HNNR_2018_rainfall.tif',
                    'NE_HNNR_2018_tas.tif',
                    'NE_HNNR_2018_tasmin.tif',
                    'NE_HNNR_2018_tasmax.tif',
                    'NE_HNNR_slope.tif'),
             mask= paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'),
             outpath = paste0(projfolder, 'NE_HumberheadNNR/Time_series_modelling/Varstacks/'),
             stackname='NE_HNNR_2018')


# modelling time series: varstack: 2015-09-30 
sat_varstack(varpath = paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),
             satimg='SEN2_20150930_lat54lon066_T30UXE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Thorne_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_HNNR_2015_rainfall.tif',
                    'NE_HNNR_2015_tas.tif',
                    'NE_HNNR_2015_tasmin.tif',
                    'NE_HNNR_2015_tasmax.tif',
                    'NE_HNNR_slope.tif'),
             mask= paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'),
             outpath = paste0(projfolder, 'NE_HumberheadNNR/Time_series_modelling/Varstacks/'),
             stackname='NE_HNNR_2015')

# modelling time series: 2016-07-19 
sat_varstack(varpath = paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),
             satimg='SEN2_20160719_lat54lon066_T30UXE_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Thorne_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_HNNR_2016_rainfall.tif',
                    'NE_HNNR_2016_tas.tif',
                    'NE_HNNR_2016_tasmin.tif',
                    'NE_HNNR_2016_tasmax.tif',
                    'NE_HNNR_slope.tif'),
             mask= paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'),
             outpath = paste0(projfolder, 'NE_HumberheadNNR/Time_series_modelling/Varstacks/'),
             stackname='NE_HNNR_2016_07')

# modelling time series: 2019-08-23
sat_varstack(varpath = paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),
             satimg='SEN2_20190823_lat54lon066_T30UXE_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Thorne_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_HNNR_mean_rainfall.tif',
                    'NE_HNNR_mean_tas.tif',
                    'NE_HNNR_mean_tasmin.tif',
                    'NE_HNNR_mean_tasmax.tif',
                    'NE_HNNR_slope.tif'),
             mask= paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'),
             outpath = paste0(projfolder, 'NE_HumberheadNNR/Time_series_modelling/Varstacks/'),
             stackname='NE_HNNR_2019')

# modelling time series: 2020-06-25
sat_varstack(varpath = paste0(projfolder,'NE_HumberheadNNR/Variable_layers/'),
             satimg='SEN2_20200625_lat54lon066_T30UXE_ORB137_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_Thorne_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('NE_HNNR_mean_rainfall.tif',
                    'NE_HNNR_mean_tas.tif',
                    'NE_HNNR_mean_tasmin.tif',
                    'NE_HNNR_mean_tasmax.tif',
                    'NE_HNNR_slope.tif'),
             mask= paste0(projfolder,'NE_HumberheadNNR/Shapefiles/Humberhead_Site_Boundaries.shp'),
             outpath = paste0(projfolder, 'NE_HumberheadNNR/Time_series_modelling/Varstacks/'),
             stackname='NE_HNNR_2020')

# finished message
sink(fs::path(projfolder, 'FINISHED.txt'))
lubridate::now()
tictoc::toc()
sink()
```

## Extracting training data

```{r,eval=F}

library(raster)
library(magrittr)
options("rgdal_show_exportToProj4_warnings"="none")
source("train_extract.R")
projfolder <- 'PeatlandCondition/'

tictoc::tic()

# Hatfield
# 2016-04-20
train_extract(varstack=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Varstacks/NE_HNNR_2016_04.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope'),
              trainr=paste0(projfolder,'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/bare_pcov_mosaic_hatfield_2016.tif'),
              outfolder=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Training_data/'),
              stratified = T)

# 2018-07-01
train_extract(varstack=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Varstacks/NE_HNNR_2018.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope'),
              trainr=paste0(projfolder,'NE_HumberheadNNR/Fine_scale_mapping/Hatfield/bare_pcov_mosaic_hatfield_2018.tif'),
              outfolder=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Training_data/'),
              stratified = T)

# Thorne

# 2016-04-20
train_extract(varstack=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Varstacks/NE_HNNR_2016_04.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope'),
              trainr=paste0(projfolder,'NE_HumberheadNNR/Fine_scale_mapping/Thorne/bare_pcov_mosaic_thorne_2016.tif'),
              outfolder=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Training_data/'),
              stratified = T)

# 2017-05-05
train_extract(varstack=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Varstacks/NE_HNNR_2017.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope'),
              trainr=paste0(projfolder,'NE_HumberheadNNR/Fine_scale_mapping/Thorne/bare_pcov_mosaic_thorne_2017.tif'),
              outfolder=paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Training_data/'),
              stratified = T)

# finished message
sink(fs::path(projfolder, 'FINISHED.txt'))
lubridate::now()
tictoc::toc()
sink()
```



## Running the time series regression models - several trials running with different test data combinations

### Thorne 2016 and 2017

```{r, eval=F}
projfolder <- 'PeatlandCondition/'

## combine data years
training1 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2016_04_thorne.csv")) %>% dplyr::select(-X)
training2 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2017_thorne.csv")) %>% dplyr::select(-X)
alltrain <- rbind(training1,training2)
write.csv(alltrain,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_thorne_2016_2017.csv"))

## look at how many per category
alltrain %>% dplyr::group_by(barecat) %>% dplyr::summarise(n=n())


## covariate analysis
varstack <- raster::stack(paste0(projfolder,'NE_HumberheadNNR/Time_series_modelling/Varstacks/probablepeat/NE_HNNR_2017.tif'))
names(varstack) <-c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope')
# run  function to see how the variables correlate
JNCCsdms::corrVars(varstack)


## train models and run predictions
RFReg(training=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_thorne_2016_2017.csv"),
      varpredict=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Varstacks/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/thorne_2016_2017/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=5000, 
      resamp=5,
      stratified=T,fillsamp=F)

# assessing change over time
dir.create(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/thorne_2016_2017/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/thorne_2016_2017/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:5],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:6],names(pred_tifs)[6]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('NE_HumberheadNNR',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/thorne_2016_2017/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})

```


### Hatfield 2018

```{r,eval=F}
projfolder <- 'PeatlandCondition/'

## combine data years
hatfield<- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2018_hatfield.csv"))

## look at how many per category
hatfield %>% dplyr::group_by(barecat) %>% dplyr::summarise(n=n())

## train models and run predictions
RFReg(training=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2018_hatfield.csv"),
      varpredict=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Varstacks/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2018/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=10000, 
      resamp=5,
      stratified=T,fillsamp=F)

# assessing change over time
dir.create(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2018/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2018/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:4],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:5],names(pred_tifs)[5]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('NE_HumberheadNNR',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2018/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})
```


### Hatfield 2016

```{r, eval=F}
projfolder <- 'PeatlandCondition/'

## combine data years
hatfield2016 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2016_04_hatfield.csv")) 
## look at how many per category
hatfield2016 %>% dplyr::group_by(barecat) %>% dplyr::summarise(n=n())

## train models and run predictions
RFReg(training=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2016_04_hatfield.csv"),
      varpredict=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Varstacks/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=10000, 
      resamp=5,
      stratified=T,fillsamp=F)

# assessing change over time
dir.create(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:5],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:6],names(pred_tifs)[6]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('NE_HumberheadNNR',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})
```


### Hatfield 2016 + 2018

```{r,eval=F}
projfolder <- 'PeatlandCondition/'

## combine data years
hatfield2016 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2016_04_hatfield.csv")) %>% dplyr::select(-X)
hatfield2018<- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2018_hatfield.csv"))%>% dplyr::select(-X)

alltrain <- rbind(hatfield2016,hatfield2018)
write.csv(alltrain,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_hatfield_2016_2018.csv"))

## look at how many per category
alltrain %>% group_by(barecat) %>% summarise(n=n())

## train models and run predictions
RFReg(training=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_hatfield_2016_2018.csv"),
      varpredict=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Varstacks/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016_2018/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=10000, 
      resamp=5,
      stratified=T,fillsamp=F)

# assessing change over time
dir.create(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016_2018/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016_2018/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:5],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:6],names(pred_tifs)[6]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('NE_HumberheadNNR',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/Hatfield_2016_2018/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})
```


### Hatfield 2016 and 2018 and Thorne 2016 and 2017

```{r,eval=F}
projfolder <- 'PeatlandCondition/'

## combine data years
thorne1 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2016_04_thorne.csv")) %>% dplyr::select(-X)
thorne2 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2017_thorne.csv")) %>% dplyr::select(-X)
hatfield1 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2016_04_hatfield.csv")) %>% dplyr::select(-X)
hatfield2 <- read.csv(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_2018_hatfield.csv")) %>% dplyr::select(-X)
alltrain <- rbind(thorne1,thorne2,hatfield1,hatfield2)
write.csv(alltrain,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_hatfield&thorne_all.csv"))

## look at how many per category
alltrain %>% dplyr::group_by(barecat) %>% dplyr::summarise(n=n())

## train models and run predictions
RFReg(training=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Training_data/NE_HNNR_hatfield&thorne_all.csv"),
      varpredict=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/Varstacks/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/hatfield&thorne_all/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=10000, 
      resamp=5,
      stratified=T,fillsamp=F)

# assessing change over time
dir.create(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/hatfield&thorne_all/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/hatfield&thorne_all/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:5],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:6],names(pred_tifs)[6]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('NE_HumberheadNNR',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"NE_HumberheadNNR/Time_series_modelling/RegressionModels/hatfield&thorne_all/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})

```






