---
title: "CUUPeat: Brecon Beacons, Wales - Broad-scale bare peat mapping"
date: "15/09/2020"
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE,echo=F}

knitr::opts_chunk$set(echo = TRUE)

# Built with r 3.6.0
library(tmap)
library(raster)
library(dplyr)
library(sf)
library(stringr)
library(gdalUtils)
library(purrr)
library(JNCCsdms)

#package dependencies
# JNCCsdms_0.1.0    stringr_1.4.0    
# purrr_0.3.4       gdalUtils_2.0.3.2
# tmap_3.0          sf_0.9-2         
# dplyr_0.8.5       raster_3.0-7     
# sp_1.4-2 

```

<div class="mycontent">

CUU Peat Project - Welsh Government - Brecon Beacons, Wales
Broad-scale bare peat mapping processing


## Site location

```{r sitemap,message=F,warning=F}

#site boundaries
site <- sf::st_read('Breacon_Site_Boundaries.shp',quiet=T)
#S2 AOI
S2 <- sf::st_read('Proposed_S2_AOI.shp',quiet=T)
#APGB AOI
APGB <- sf::st_read('Proposed_APGB_AOI.shp',quiet=T)

#plot
tmap::tmap_mode("view")

tmap::tm_shape(site) +  tmap::tm_borders(alpha=0.5) +  tmap::tm_fill("green") +
  tmap::tm_shape(S2) +  tmap::tm_borders(alpha=1,col='blue') +
  tmap::tm_shape(APGB) +  tmap::tm_borders(alpha=1,col='pink')


```


## Preparing Variable layers

### DTM and slope

```{r dtmslopevars, eval = FALSE}

projfolder <- 'PeatlandCondition/'

#crop dtm to AOI
dtm <- raster::raster('UK_DEM_Composite_10m_WGS84.tif')
AOI= sf::st_read(paste0(projfolder,'Data/Site_AOIs/WG_Beacons/Proposed_S2_AOI.shp'),quiet=T)

# reproject AOI
AOI_WGS84 <- AOI %>% sf::st_transform(4326)

#crop to dtm
dtm_crop <-raster::crop(dtm,AOI_WGS84)
dtm_bng <- raster::projectRaster(dtm_crop, crs=sp::CRS('+init=epsg:27700'), method='bilinear')
raster::writeRaster(dtm_bng,paste0(projfolder,'WG_BreconBeacons/Variable_layers/WG_BB_dtm.tif'),overwrite=T)

# we can then create a slope layer using gdalutils
gdalUtils::gdal_setInstallation()
gdalUtils::gdaldem(mode="slope",
                     input_dem=paste0(projfolder,'WG_BreconBeacons/Variable_layers/WG_BB_dtm.tif'),
                     output=paste0(projfolder,'WG_BreconBeacons/Variable_layers/WG_BB_slope.tif'),
                    ,verbose=T)

```

### Climate layers

```{r prepclimate,eval=F}
#list all data files from met office
met_files <- data.frame(files=list.files("MetOffice_HadUK_monthly_grid_1km/",pattern='.nc'))
#subset to seasonal records
met_season <- met_files %>% dplyr::filter(stringr::str_detect(files, "Seas"))
#iterate through taking just the summer season data
out<- purrr:::map(met_season$files,function(metlayer){
  met_raw <- raster::brick(paste0("MetOffice_HadUK_monthly_grid_1km/",as.character(metlayer)))
  summer <- met_raw[[3]]
  summer_crop <-raster::crop(summer,AOI)
  #get filename of measure and year
  filename_split <- stringr::str_split(gsub(as.character(metlayer),pattern="Had",replacement="_Had"),pattern='_')
filename <- paste0(stringr::str_extract(as.character(filename_split[[1]][4]),'\\d{4}'),'_',filename_split[[1]][2])
#write out
  raster::writeRaster(summer_crop, paste0(projfolder,'WG_BreconBeacons/Variable_layers/WG_BB_',filename,'.tif'),overwrite=T)
  summer_crop
})


#create a mean layer for use with years without data released - 2019,2020
vars <- c("rainfall","tas","tasmin","tasmax")
# iterate through each variable
all <- purrr::map(vars, .f=function(var){
  varfiles <- list.files(paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),full.name=T,pattern=var) # list all files in that folder for variable
  varfiles_r <- purrr::map(varfiles, raster::raster) #load
  varstack <- raster::stack(varfiles_r) #stack 
  varmean <- raster::calc(varstack,mean) #mean
  raster::writeRaster(varmean,paste0(projfolder,'WG_BreconBeacons/Variable_layers/WG_BB_mean_',var,'.tif')) #write out
})

```



### Sentinel imagery and calculating indices

extraction time series: 
* S2A_20170525_lat52lon366_T30UVC_ORB037_utm30n_osgb	2017-05-25
* S2B_20180505_lat52lon366_T30UVC_ORB037_utm30n_osgb	2018-05-05


Modelling time series:
* 2016	S2A_20160719_lat52lon366_T30UVC_ORB037_utm30n_osgb	2016-07-19
* 2017	S2A_20170525_lat52lon366_T30UVC_ORB037_utm30n_osgb	2017-05-25
* 2018	S2A_20180629_lat52lon366_T30UVC_ORB037_utm30n_osgb	2018-06-29
* 2019	S2B_20190420_lat52lon366_T30UVC_ORB037_utm30n_osgb	2019-04-20
* 2020	S2B_20200623_lat52lon366_T30UVC_ORB037_utm30n_osgb	2020-06-23


```{r, eval=F}
projfolder <- 'PeatlandCondition/'

#prep and crop sentinel imagery to new AOI
s2_processing(s2path=paste0(projfolder,"Data/Sentinel-2/WG_BB/"),
              out_folder = paste0(projfolder,"WG_BreconBeacons/Variable_layers/"),
              nirband=8,rband=3,bband=1,gband=2,swirband=9,
              indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
              )
```


## Compiling Variable Stacks

```{r,eval=F}
# create varstacks and mask

# do this for the highly probable peat crop as well as the probable peat

### highly probable peat ###
# extraction varstack: 2017-05-25 (same as 2017 time series)
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20170525_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2017_rainfall.tif',
                    'WG_BB_2017_tas.tif',
                    'WG_BB_2017_tasmin.tif',
                    'WG_BB_2017_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp)',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2017')
# extraction varstack: 2018-05-052018-05-05
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20180505_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2018_rainfall.tif',
                    'WG_BB_2018_tas.tif',
                    'WG_BB_2018_tasmin.tif',
                    'WG_BB_2018_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2018_extract')
# 2016 varstack: 2016-07-19
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20160719_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2016_rainfall.tif',
                    'WG_BB_2016_tas.tif',
                    'WG_BB_2016_tasmin.tif',
                    'WG_BB_2016_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2016')
# 2018 varstack: 2018-06-29
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20180629_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2018_rainfall.tif',
                    'WG_BB_2018_tas.tif',
                    'WG_BB_2018_tasmin.tif',
                    'WG_BB_2018_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2018')
# 2019 varstack: 2019-04-20
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20190420_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_mean_rainfall.tif',
                    'WG_BB_mean_tas.tif',
                    'WG_BB_mean_tasmin.tif',
                    'WG_BB_mean_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2019')
# 2020 varstack: 2020-06-23
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20200623_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_mean_rainfall.tif',
                    'WG_BB_mean_tas.tif',
                    'WG_BB_mean_tasmin.tif',
                    'WG_BB_mean_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2020')

### probable peat ###
# extraction varstack: 2017-05-25 (same as 2017 time series)
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20170525_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2017_rainfall.tif',
                    'WG_BB_2017_tas.tif',
                    'WG_BB_2017_tasmin.tif',
                    'WG_BB_2017_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2017')
# extraction varstack: 2018-05-052018-05-05
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20180505_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2018_rainfall.tif',
                    'WG_BB_2018_tas.tif',
                    'WG_BB_2018_tasmin.tif',
                    'WG_BB_2018_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2018_extract')
# 2016 varstack: 2016-07-19
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20160719_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2016_rainfall.tif',
                    'WG_BB_2016_tas.tif',
                    'WG_BB_2016_tasmin.tif',
                    'WG_BB_2016_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2016')
# 2018 varstack: 2018-06-29
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20180629_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2018_rainfall.tif',
                    'WG_BB_2018_tas.tif',
                    'WG_BB_2018_tasmin.tif',
                    'WG_BB_2018_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2018')
# 2019 varstack: 2019-04-20
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20190420_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_mean_rainfall.tif',
                    'WG_BB_mean_tas.tif',
                    'WG_BB_mean_tasmin.tif',
                    'WG_BB_mean_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2019')
# 2020 varstack: 2020-06-23
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20200623_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_mean_rainfall.tif',
                    'WG_BB_mean_tas.tif',
                    'WG_BB_mean_tasmin.tif',
                    'WG_BB_mean_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'fine_peat_soils.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/'),
             stackname='WG_BreconBeacons_2020')


#### probable peat###
# extraction varstack: 2017-05-25 (same as 2017 time series)
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20170525_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2017_rainfall.tif',
                    'WG_BB_2017_tas.tif',
                    'WG_BB_2017_tasmin.tif',
                    'WG_BB_2017_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'CUU_WG_Peat_Area_of_Interest_Data.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/'),
             stackname='WG_BreconBeacons_2017')
# extraction varstack: 2018-05-052018-05-05
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20180505_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2018_rainfall.tif',
                    'WG_BB_2018_tas.tif',
                    'WG_BB_2018_tasmin.tif',
                    'WG_BB_2018_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'CUU_WG_Peat_Area_of_Interest_Data.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/'),
             stackname='WG_BreconBeacons_2018_extract')
# 2016 varstack: 2016-07-19
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20160719_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2016_rainfall.tif',
                    'WG_BB_2016_tas.tif',
                    'WG_BB_2016_tasmin.tif',
                    'WG_BB_2016_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'CUU_WG_Peat_Area_of_Interest_Data.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/'),
             stackname='WG_BreconBeacons_2016')
# 2018 varstack: 2018-06-29
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20180629_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_2018_rainfall.tif',
                    'WG_BB_2018_tas.tif',
                    'WG_BB_2018_tasmin.tif',
                    'WG_BB_2018_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'CUU_WG_Peat_Area_of_Interest_Data.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/'),
             stackname='WG_BreconBeacons_2018')
# 2019 varstack: 2019-04-20
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20190420_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_mean_rainfall.tif',
                    'WG_BB_mean_tas.tif',
                    'WG_BB_mean_tasmin.tif',
                    'WG_BB_mean_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'CUU_WG_Peat_Area_of_Interest_Data.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/'),
             stackname='WG_BreconBeacons_2019')
# 2020 varstack: 2020-06-23
#varstacks:"Brightness","NDVI","RG","RB","NBR","NDWI","GLI", nir band, average seasonal temp, max seasonal temp, precip, slope
sat_varstack(varpath = paste0(projfolder,'WG_BreconBeacons/Variable_layers/'),
             satimg='SEN2_20200623_lat52lon366_T30UVC_ORB037_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_BreconBeacons_msk.tif',
             satbands=8,
             indices=c("Brightness","NDVI","RG","RB","NBR","NDWI","GLI"),
             vars=c('WG_BB_mean_rainfall.tif',
                    'WG_BB_mean_tas.tif',
                    'WG_BB_mean_tasmin.tif',
                    'WG_BB_mean_tasmax.tif',
                    'WG_BB_slope.tif'),
             mask= 'CUU_WG_Peat_Area_of_Interest_Data.shp',
             outpath = paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/'),
             stackname='WG_BreconBeacons_2020')

```

## Extracting training data

```{r,eval=F}
# Two dates needed: 2017 and 2018 
# 2017-05-25
train_extract(varstack=paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/WG_BreconBeacons_2017.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope'),
              trainr=paste0(projfolder,'APGB/WG_BB/Combined/bare_pcov_mosaic_2017.tif'),
              outfolder=paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Training_data/'),
              stratified = T)


# 2018-05-05
train_extract(varstack=paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/highlyprobablepeat/WG_BreconBeacons_2018_extract.tif'),
              varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope'),
              trainr=paste0(projfolder,'Data/APGB/WG_BB/Combined/bare_pcov_mosaic_2018.tif'),
              outfolder=paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Training_data/'),
              stratified = T)

```

## Running the time series regression models

```{r,eval=F}
projfolder <- 'PeatlandCondition/'

## combine data years
training1 <- read.csv(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2017.csv")) %>% dplyr::select(-X)
training2 <- read.csv(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2018_extract.csv")) %>% dplyr::select(-X)
alltrain <- rbind(training1,training2)
write.csv(alltrain,paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2017_2018.csv"))

## look at how many per category
alltrain %>% group_by(barecat) %>% summarise(n=n())


## covariate analysis
varstack <- raster::stack(paste0(projfolder,'WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/WG_BreconBeacons_2017.tif'))
names(varstack) <-c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",'slope')
# run  function to see how the variables correlate
JNCCsdms::corrVars(varstack)


## train models and run predictions
RFReg(training=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2017_2018.csv"),
      varpredict=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017 & 2018/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=5000, 
      resamp=5,
      stratified=T,fillsamp=F)

```



## Assessing change over time


```{r,eval=F}
dir.create(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017 & 2018/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017 & 2018/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:4],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:5],names(pred_tifs)[5]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('WG_BreconBeacons',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017 & 2018/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})

```


# rerunning on just 2017 data with 5000 points

```{r,eval=F}
## Running the time series regression models
projfolder <- 'PeatlandCondition/'

## combine data years
training <- read.csv(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2017.csv")) 
## look at how many per category
training %>% group_by(barecat) %>% summarise(n=n())

## train models and run predictions
RFReg(training=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2017.csv"),
      varpredict=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=5000, 
      resamp=5,
      stratified=T,fillsamp=F)

# running the change analysis
dir.create(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:4],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:5],names(pred_tifs)[5]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('WG_BreconBeacons',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})
```



# rerunning on just 2017 data with 2000 points

```{r,eval=F}
## Running the time series regression models
## combine data years
training <- read.csv(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2017.csv")) 
## look at how many per category
training %>% group_by(barecat) %>% summarise(n=n())

## train models and run predictions
RFReg(training=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Training_data/WG_BreconBeacons_2017.csv"),
      varpredict=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/Varstacks/probablepeat/"),
      varnames=c("NIRBand","Brightness","NDVI","RG","RB","NBR","NDWI","GLI","rainfall","av_temp","min_temp","max_temp",
                    'slope'),
      out.folder=paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017_2000points/"),
      max_tries=10,
      prop.test=0.25,
      nsamp=2000, 
      resamp=5,
      stratified=T,fillsamp=F)

# running the change analysis
dir.create(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017_2000points/ChangeMaps"))

pred_maps <- list.files(paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017_2000points/Outputs/"),
                        pattern='.tif',full.names = T)
pred_tifs <- purrr::map(pred_maps,raster::raster)

names(pred_tifs) <- gsub('.tif',basename(pred_maps),replacement="")

year_combo <- data.frame(startyear=c(names(pred_tifs)[1:4],names(pred_tifs)[1]),endyear=c(names(pred_tifs)[2:5],names(pred_tifs)[5]))

purrr::map2(as.character(year_combo$startyear),as.character(year_combo$endyear),.f=function(x,y){
  start_bare <- pred_tifs[[x]]
  end_bare <- pred_tifs[[y]]
  change_bare <-raster::overlay(start_bare,end_bare,fun=function(x,y){y-x}) 
  endyear <- gsub('_barepeat',gsub('WG_BreconBeacons',y,replacement=''),replacement='')
  raster::writeRaster(change_bare,paste0(projfolder,"WG_BreconBeacons/Time_series_modelling/RegressionModels/2017_2000points/ChangeMaps/",x,endyear,'change.tif'),overwrite=T)
})
```



